### 2015-05-10 thesis/thesis/chap4/chap4.tex
rabeb comments
added coop

\chapter{Training Chapter}\label{cha:PCTLKC}

@@

Over the past years, online services have become an important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One practical solution towards delivering high quality services is utilizing intelligent decision making agents. These agents aim at maximizing their gain by exploring the best ways to provide services that satisfy end users\cite{Zeng:2003:QDW:775152.775211, 10.1109/ARES.2008.7, Demirkan2013412, journals/tsc/ZhengZYB13, Josang:2007:STR:1225318.1225716}. However, agent-based web services are functionally limited in the sense that they cannot handle a large number of requests at the same time without compromising the quality of service provided. Recent developments have attempted to shift web services from passive models, consisting of individual components, to models made up of autonomous and group-based components that share common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that directly impact the group's overall performance in achieving common goals. To that end, we see the emergence of web service \emph{communities}, which consist of services with similar functionalities but distinct nonfunctional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web services runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

The advantages of community of web services are that the community facilitates the discovery of web services and provides better quality of service compared to passive models with individual components. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. But, they regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined work of a number of web services. The downside to communities is the complexity of management involved in finding and inviting individual web services and managing the overall quality of the combined work of several web services.
% because although they have similar functionality, they have different attitudes.
When interacting with communities of web services, users send their requests to the coordinators of the community, which play the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks web services that could harm the value of the community.

\textbf{Challenge and Problem Statement.} In recent work, communities of web services have been proposed in order to facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. However, there is a twofold challenge that yet to be addressed. The first part is community development, which involves finding the right web services for the community, satisfying the objectives of all the web services and community coordinators involved within the community; and the second part is from the individual web service's side because each web service aims to join the community that would guarantee the best outcome. The advocated solutions \cite{conf/webist/MaamarLBTS07, 10.1109/ARES.2008.7, journals/soca/XuYLZB11, 10.1109/TSC.2012.12, managing-hela-jalel, DBLP:conf/IEEEscc/KhosravifarABT11, DBLP:conf/IEEEscc/LimTMB12} have attempted to address these research problems. These research problems are reflected by the two challenge parts:
\begin{enumerate}
	\item The solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. The problem is that in real world scenarios, decisions made by independent service providers are highly distributed.
	\item The solutions either propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy to follow, or oversimplify the problem by eliminating important parameters and using approximation techniques to make the algorithms tractable \cite{10.1109/TSC.2012.12}.
\end{enumerate}
These approximation methods sometimes negatively influence the outcome because simplifying the constraints may cause important aspects of the problem to be ignored.

For example, instead of calculating the gain distribution using the complex shapely-value method, \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores the gain generated from collaborative work of sub-communities. Other categories of related work exist that do not focus on all the parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}.
%And some other work which do not focus on rationality on web services or communities involved [refrences].
In \cite{journal-community-formation} we proposed a cooperative game theory based model for aggregating web services in communities. A centralized decision maker in communities based on a complete knowledge of available web service quality metrics and performance, has been used to form optimal and stable communities that maximize individual and group income. However, centrality and complete information are strong assumptions, which are not very compatible with real business scenarios.

\textbf{Contributions.} In this paper, we introduce DDM, a Distributed Decision Making model for  community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree is created from the past data and reflects comprehensive vision about agents' attitude of action selection. However, the tree is subject to get continuously modified to adapt with environmental changes.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services to efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in decision making process and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. In other words, the web services agents are equipped with a recommender system to suggest beneficial action. Moreover, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their short-term and long-term goals; therefore, opting for best decision available.

To effectively generate the decision model for web services, we used real data set to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. The data set has been extracted by real-world QoS evaluation results from 142 users on 4,532 Web services on 64 different time slots. Combining these data based on each web service point of view on different time slots, we acquired 5 different unique features for 4,532 web services. By engineering and extracting these features, we gathered functional and cooperative features for both individual web services and communities in different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in the training model. Using cross validation sets, web services are able to compute the optimal hypothesis function (using logistic regression) that can be used to predict outcomes of cooperative work with other individual web services or communities. Our findings show that web services equipped with DDM have by far better outcomes in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

\section{Preliminaries and Challenging Issues}\label{s:preliminaries}
%In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.
In this section, we discuss the preliminary concepts of communities of web service and we introduce the challenges of community formation.

\subsection{Web Services}\label{s:ws}

In recent years, online services have become a standard part of daily life around the globe. Many modern applications rely on web services from different providers. For instance, many mobile and tablet applications that have limited storage and processing power are merely aggregating information from different online services. Examples are vast, including weather forecasting, ticket selling, shopping apps, local maps and location services.

The World Wide Web Consortium (W3C) defines web services as ``software systems designed to support interpretable machine-to-machine interaction over a network. It has an interface described in a machine-processable format (specifically WSDL). Other systems interact with the web service in a manner prescribed by its description using SOAP messages, typically conveyed using HTTP with XML serialization in conjunction with other Web-related standards.'' When developers declare a new web service, it will be
discovered based on its description, which fully discloses its functionalities. Developers also have to declare a public interface and a readable documentation to help other developers when integrating different services \cite{w3cwsdl}. Nowadays, web API, standards that do not require XML-based web service protocols like SOAP and WSDL are also emerging. They are called REST (representational state transfer) services, which are moving towards simpler communication protocols.
%They are not restricted
%to XML formats, recently JSON, a human readable and simpler format
%is becoming popular among online service providers.

We are not going to delve into the engineering details of online web service implementation and its protocols in this paper. We are interested in web services from a business model perspective. Service providers usually charge end users for services they provide. For example, Google has listed pricing and plans for a wide range of services they provide on their web service console page\footnote{https://code.google.com/apis/console}.

In this paper, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} that provide services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenues, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to enhance efficiency by collaborating with others, to have access to broad market share, and for the opportunity to receive a bigger task pool from end users. Furthermore, the high reliance on web services has resulted in increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

\subsection{Web Service Communities}\label{s:wsc}

The community of web services is essentially a virtual group of web services having similar and complementary functionalities\cite{DBLP:journals/ijebr/MaamarSTBB09}. Communities aggregate web services and communicate with other entities such as UDDI registries and users, using identical protocols to those used by single web services. Web services join communities to increase utility by having a larger market share and task pool. Community coordinators are responsible for community development, managing membership requests from web services and distributing user tasks among the community members. Community coordinators try to attract quality web services to join and keep the community as stable and productive as possible to gain a better reputation and user satisfaction, which increases the community's market share. How web services reside within communities and how communities of web services are engineered is described comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\subsection{The Join Challenge}\label{s:tjc}
Sometimes, web services can increase their overall utility by collaborating with other web services as communities. This collaboration helps them with better ways of sharing resources and having higher reputation, better market share and wider visibility. Web services and communities come with different internal and external parameters, and the long-term outcome is dependent on these parameters.

The goal of all parties involved in the community is to maximize their long-term outcome while they are operating as part of the community. Web services need to be equipped with a selection strategy to choose from all of the different possible collaboration groups they can form as well as an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment with different possible collaborative groups and other web services in order to estimate their gain over time. However, with a high number of possible communities, it is not possible to test collaboration with random web services. Even if a linear approximate function for estimating utility based on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of the possible number of partitions of web services into communities would make any brute-force type algorithm for the best community selection strategy intractable and impractical in real-world application settings.

\subsection{Join Consequences}\label{s:jc}
It is worth mentioning that a \emph{join} event takes place as a result of two parties that are looking to expand their collaborations. All actions are chosen in an attempt to enhance the overall outcome. However, the selected action may result in decreasing the overall utility in the long run.
% long term.
This is the case when a single web service joins a community, but the complex process of task allocation eliminates the visibility of the web service and, thus, the web service stays idle within the community, which makes the join action a bad decision. The same event might be beneficial for the community, as it hosts a new web service that can engage in performing a new coming task. But overall, neither side benefits from collaborating with the other and the join event has negative consequences for at least one side's utility.

The more common scenario is when both parties benefit from the joining of a web service to a community. This action is rational, as both the web service and the community enhance the utility. However, the community may not be the best choice for the web service to join. In other words, the web service could have joined a better community, but, due to of lack of knowledge about the surrounding environment, joined the available community which was not the optimal choice. Since the community does enhance its utility, the web service could stay with that community, which results in a sub-optimal increase in web service's utility. In the following section the proposed model provides solutions that effectively address the aforementioned challenges.

\section{The Model Components}\label{s:themodelcomponents}

In this section, we discuss the parameters that we use in the rest of the paper. Then we present the task distribution and revenue model of our distributed web services communities.

\subsection{Internal Features}\label{s:if}

With a group of web services having identical or similar functionalities, QoS metrics provide nonfunctional characteristics for optimal candidate selection. Web service quality metrics were presented in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}. We have adopted the most representative QoS properties of web services that highly influence the utility of web services.
%We refer to a typical web service as ws_{i}.

Let $C = \{ws_1,ws_2,..., ws_n\}$ be a community with $|n|$ web services. We define the following features for the group of web services based on their functional parameters:

\begin{itemize}

  \item \emph{Throughput $(Th_{C})$} is the rate at which a service can process requests. QoS measures can include the maximum throughput or a function that describes how throughput varies with load intensity. Throughput is a positive real number. For communities the expected throughput value of collaborating web services can be estimated as summation of throughputs of all the services inside:

	\begin{equation}
		 Th_{C} = \left\{
			\begin{array}{l l}
				Th_{w} & \quad \text{if $|C| = 1$}\\
				\sum_{w \in C}{(Th_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	\item \emph{Availability $(A_{C})$} is the percentage of time that a service is operating.
	The probability that the service operation is accessible. Availability is also a real number in the range $[0, 1]$. For communities the expected availability of collaborating web services considering they follow a parallel system can be estimated as:

	\begin{equation}
		A_{C} = \left\{
			\begin{array}{l l}
				A_{w} & \quad \text{if $|C| = 1$}\\
				1-\prod_{w \in C}{(1-A_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	\item \emph{Execution Time $(Et_{C})$} is the time a service takes to respond to various types of requests.
	%is the expected delay between the time instant when a request is sent and the time when the result is obtained.
	Execution time is measured in milliseconds. Execution time can be affected by load intensity, which can be measured in terms of arrival rates (such as requests per second) or number of concurrent requests. This internal feature is a positive integer. For communities, the expected execution time of collaborating web services can be estimated as the execution time of the bottleneck service which is the service with the slowest execution time:

	\begin{equation}
		Et_{C} = \left\{
			\begin{array}{l l}
				Et_{w} & \quad \text{if $|C| = 1$}\\
				max_{w \in C}{(Et_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	%\item \emph{Data Quality} The ability of a data collection to meet user requirements , defined as the proximity of a value v returned by web service to a value considered as correct. The measure of data quality is considered here as a real number in the range [0, 1], where 1 represents the most desirable score.
\end{itemize}

	We normalize the range of these features so that each feature contributes approximately proportionately to the final utility outcome value. We adopt the \emph{standardization} method. We subtract the \emph{mean} from each feature then we divide the values of each feature by its \emph{standard deviation}.

%$HI = \overline{HI}$

\subsection{External Features}\label{s:ef}

The quantitative values of quality metrics need some benchmark values to represent their goodness. In other words, without some benchmark values it would be difficult for web services to identify their goals, and performance quality at any specific value of these metrics. Therefore, we have introduced two external features for assessing web services' estimate with regard to their standing among other web services.

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of all the other web services, about which the web service has information of them. The less value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of all the other web services,, about which the web service has information of them. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

A community of web services is abstracted as a standard web service that provides services that share the same internal and external features mentioned for web services. Inside, they adopt different protocols to manage the community; however, other web services or end users can communicate with communities of web services, using standard protocols, and inner details will be encapsulated for their clients.

\subsection{Task Distribution}

Communities of web services usually employ an implementation of Contract-Net protocol for task distribution, in which services opt for incoming tasks, and receive the tasks for which they ask. However, in our model, our community members would try to distribute tasks based on their capabilities and the QoS parameters provided by the web services. We have used a slightly modified \emph{weighted fair queuing} method to distribute tasks among community members. The goal is to allocate incoming tasks to web services with a rate matching the throughput value of $Th_{ws}$. In the \emph{weighted fair queuing} method, the input flow is multiplexed along different paths; however, in our case, if the rate of incoming tasks is less than the community's total throughput $(Th_{C})$, which is the summation of throughput values of the web services in the community, some of the input tasks will be queued and served with a delay.
%Thus, the amount of tasks performed by the community is $\sum_{ws}{Th_{ws}}$ when $\sum_{ws}{Th_{ws}} \leq R_{C}$.
However, when the incoming task rate is less than the throughput of the community, the \emph{weighted fair queuing} algorithm assigns a weighted task rate of $|input Task Rate| \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each web service ($ws$) within the community.

While distributing tasks, the community members can verify the performance, throughput and quality of service of tasks being performed by web services. The community can assess if those web services are capable of performing the number of tasks they advertised. If for any reason, there is a decline in the quality metric or throughput, the community can consider the new values as a benchmark for future performance calculations, and penalize the suspicious web services. This way, players will have incentive to truthfully disclose their actual capabilities in order to maximize profit from the community and to avoid being penalized. In addition, the system should be dynamic enough to detect and react to web services' quality metrics variation, as over time, web service metrics may degrade or improve, changes to which the community should adjust.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

\subsection{Community Revenue}

>>>>

\chapter{Coopetitive Behavior of Services within Communities}\label{cha:PCTLKC}

@@

In previous chapter, our focus was on community formation and we emphasised on cooperative behavior of the web services as agents. Within communities, those services, selfish and utility maximizers by nature, can follow two different strategies, namely cooperation and competition in order to increase their payoffs when they provide services to consumers \cite{VuFind-10008938119}. In typical business settings, services are used to compete within communities as they provide the same functionalities and the number of users requests is finite. However, the same reason of providing similar functionalities can lead services to cooperate because they can replace each other in case of failure or unavailability, and services can do better in a coalition structure. Analyzing services competition and cooperation strategies within communities is still an open problem that motivates the research described in this section. we propose a mechanism within which service agents
in the community could choose either to compete for an announced task\footnote{Requests and tasks are used in this report interchangeably.}, or to cooperate with other competing services in the same community to accomplish some subtasks of the announced task. We equip intelligent web services to follow a reasoning technique to choose best interactive strategy (Coopetitive attitude, which is categorized to compete and cooperate). In the proposed system, We explore details behind the strategic decision making procedures and enable service agents to apply different techniques to constrain high efficiency and obtain the maximum utility. We investigate services' expected payoffs and the involved probabilities that are used to choose over the two interacting strategies.

%\subsection{The Proposed Framework}\label{The Proposed Framework}
Here, we first present the architecture of the proposed
model. We explore the characteristics of intelligent service
agents and their network. We link this architecture to the
implemented system where we investigate the services' coopetitive
attitudes. We compute the involved system parameters and explain
the services' interactive strategy profiles by highlighting their
coopetitive choices.

\begin{figure}[h]
\centering
\includegraphics[scale=1]{Figures/architecture++.eps}
\caption{Services are partitioned into competitive and cooperative
sets. Competitive services may get tasks directly from the master
agent and they can share it with other cooperative services in
their collaborative networks within the same community.}
\label{architectureFigure}
\end{figure}

\section{Architecture}

Figure \ref{architectureFigure} illustrates the architecture of a
typical community aggregating a number of services with different
interactive strategies. Some of them compete for the task where
they directly deal with the master. Some others cooperate in the
associated task where they only deal with the competed service as
the task leader and do not directly interact with the master (the
master deals only with the service that has bid for the task,
which is responsible of choosing its collaborative network). In
both sets, some service agents are for certain moments out of any
collaboration network. Upon allocation of
the task, the service is responsible for offering the required QoS
that is stated in the task being generated by a consumer.
Afterwards,  the master rewards or penalizes the competing
service by upgrading or degrading its reputation according to the
offered QoS compared with the required one. This comparison
influences the sorting mechanism used by the master to allocate
the tasks in further task allocation rounds.

%ok til here

\section{System Parameters}
In this section, we demonstrate the involved parameters and their
corresponding formulations and explanations.

\textbf{Task QoS} ($T_{QoS}^r$) is the required QoS metric for a
specific task $r$. Users define tasks with specific QoS
requirements such as response time, availability, and
successability (or accuracy). We aggregate and
normalize these metrics to a value between $0$
and $1$. %Throughout this paper, we refer to this value as
%$T_{QoS}$.

\textbf{Service QoS} ($QoS_w^r$) is the QoS provided by the
service $w$ after performing the task $r$. Again, the metrics that
contribute in computing this QoS are aggregated and normalized to
a value between $0$ and $1$. The offered quality might or might
not meet the required task quality $T_{QoS}^r$. In the latter
case, the service user would be disappointed and a negative
satisfaction feedback is expected. In our proposed system, both
cases are considered when calculating the services' reputation.

%\textbf{NTD} is the number of tasks completed by a web service.

\textbf{Budget} ($B_w^t$) is the amount of money the service agent
$w$ has in its disposal during the window time $t$ (i.e.,
$[0,t]$), which helps pay for the community membership fees
($\epsilon$) and is one of the parameters that the service agent
considers when deciding about getting involved in a competition or
not.

\textbf{Reputation} ($Rep_w^t$) is a significant factor in any
online community \cite{Fouss:2010:PRM:1751668.1751727}. Without a reputation enabling
mechanism, users cannot differentiate among services, specially
the ones which offer the same type of service. Reputation
mechanisms usually aggregate users' experiences and in our case it
strongly depends on QoS that each service provides. Users define
tasks, each one with specific quality $T_{QoS}^r$, so that after
performing a certain number of tasks, each one with $QoS_w^r$,
during a window time $t$, the reputation of $w$ gets evaluated by
the master agent. $Rep_{w}^{t}$ refers to the reputation of $w$
during that window time $t$.

In Equation \ref{repr}, we compute the reward that the master
computes considering the task $r$'s QoS $T_{QoS}^r$ compared with
the service offered quality $QoS_w^r$. In case the offered quality
meets user expectations, the reward value would be positive. In
this system, we consider a small value as default rewards $\eta$
which the master considers together with the proportional level of
satisfaction as a weighted value (by $\upsilon$). In this case,
the higher the offered quality, the more weighted reward. In case
the offered quality did not meet the user expectations, the reward
would be negative. A default penalty value $\rho$ (where
$\rho>\eta$) together with the weighted proportional difference
are therefore considered. The idea is to harshly penalize the
services rather than rewarding them. To this end, rational service
agents should carefully analyze their capabilities once the
available tasks are announced. Equation \ref{reward} computes the
obtained reward by $w$ during the window time $t$ considering the
set $task_w^t$ of tasks performed by $w$ during the window time
$t$. In our proposal, service agents have the goal of increasing
their budget, which is directly related to their reputation. Thus,
they have to decide strategically how to maximize this value.

\begin{equation} \label{repr}
reward_w^r = \begin{cases}
\eta + \upsilon \frac{QoS_w^r}{T_{QoS}^r+QoS_w^r}   & \text{if $T_{QoS}^r\leq QoS_w^r$;}\\
-(\rho +  \upsilon \frac{T_{QoS}^r}{T_{QoS}^r+QoS_w^r} ) & \text{otherwise.}\\
\end{cases}
\end{equation}

\begin{equation} \label{reward}
reward_w^t =
\begin{cases}
\frac{\sum_{r \in
task_w^t}reward_w^r}{|task_w^t|}& \text{if $task_w^t \neq \emptyset$;}\\
0 & \text{otherwise.}\\
\end{cases}
\end{equation}

The assigned reputation value is updated by the computed reward
value. The computed reputation of services is bounded by the
minimum and maximum reputation values $0$ and $1$. % to constrain
%the balance the cooperative community of web services. The updated
Let $\Gamma = Rep_{w}^{t} + reward_w^t$. The updated reputation
value is then computed as follows:

\begin{equation}\label{repz}
%Rep_{w}^{t+1} = Rep_{w}^{t} + reward
%\end{equation}
%\begin{equation*}
Rep_{w}^{t+1} = \begin{cases}

\Gamma & \text{if $ 0 \leq \Gamma \leq 1$;}\\
0  & \text{if $\Gamma < 0$;}\\
1 & \text{if $\Gamma > 1$.}\\
\end{cases}
\end{equation}
%
For new services with no previous reputation value, we use the
bootstrapping trust technique proposed in \cite{Yahyaoui2}. This
technique consists in giving the new services a chance and observe
their behaviors for a period of testing time. The observation
sequence is modeled as a hidden Markov model that is used to
detect the behavior of the service by comparing the observation
behavior against pre-defined trust patterns. Based on the matching
result, an initial value is assigned to the service. Using this
initial reputation value, services quickly converge to their
actual and stable values using the update function.

\begin{proposition}\label{Complexity-Rep}
$Rep_{w}^{t}$ can be computed in time $O(|t|)$, i.e., in time
linear in the size of the window $t$.
\end{proposition}

\begin{proof}\footnote{In this paper, we
assume that the common arithmetic and elementary functions, such
as multiplication, division and trigonometric functions can be
computed in time $O(1)$ as they operate on inputs of fixed sizes.}
The function $Rep_{w}^{t}$ is recursive on $t$, but the algorithm
works by storing the last calculated reputation value in a
variable, so it will not be recalculated again at each iteration.
However, the calculation of $reward_w^t$ is needed. Since the
function $reward_w^t$ can be computed in time linear in the number
of tasks (see Equations \ref{repr} and \ref{reward}), which in
turn is linear in the size of the window time, the result follows.
$\Box$
\end{proof}

\textbf{Growth Factor} ($G_w^t$) is a parameter which declares
services' performance based on their recent strategies and
activities. Growth factor is relative to services' reputation
$Rep_w^t$, QoS during the window time $t$ $QoS_w^t$, and budget
$B_w^t$. This factor is the main variable a typical service uses
to decide about which strategy to adopt. We use Equation \ref{eq:growthfactor} to compute the
growth factor $G^t_w$ of the service $w$ during the window time
$t$ as the average of the three aforementioned parameters, where
$n_t$ is the total number of offered tasks to the whole community
during the
window time $t$, %(we consider the time window $t$ large enough so
%that $n_t \neq 0$),
$\mu_w$ is the mean received service fee, and $\epsilon$ is the
cost of community membership.

\begin{equation}\label{eq:growthfactor}
G^t_w = \frac{Rep^t_w + QoS_w^t+\frac{B_w^t}{n_t  \mu_w -
\epsilon}}{3}
\end{equation}
\begin{equation*}
\mu_{w} \in\{\mu_{w, CM}, \mu_{w, CO}\},~~~QoS_w^t =
\begin{cases} \frac{\sum_{r \in
task_w^t}QoS_w^r}{|task_w^t|}& \text{if $task_w^t \neq \emptyset$;}\\
0 & \text{otherwise.}\\
\end{cases}
\end{equation*}

This equation is designed so that it satisfies the following
desirable properties:

\begin{enumerate}

\item The growth factor function should be monotonically
increasing in the offered quality of service $QoS_w^t$.

\item The growth factor function should be monotonically
increasing in the service's reputation $Rep^t_w$.

\item The growth factor function should be monotonically
increasing in the budget $B_w^t$ if the maximum possible profit is
positive and monotonically decreasing in $B_w^t$ if the maximum
possible profit is negative. This property reflects the idea that
the budget contributes in the increase of the growth factor as far
as there is a chance to make profit. In fact, the contribution of
the budget $B_w^t$ in the calculation of the growth factor should
be proportional to the maximum possible profit $n_t \mu_w -
\epsilon$ where the service $w$ receives all the offered tasks
during the window time $t$.

\end{enumerate}

  % if the profit $\beta_w$ made by the web service $w$
%considering the mean received service fee $\mu_w$ and the cost of
%community membership $\epsilon$ is positive.
It is easy to show that Equation \ref{eq:growthfactor} satisfies
the three aforementioned properties by calculating the partial
derivatives $\partial G^t_w$ of this function in 1) $QoS_w^t$
($\frac{\partial G^t_w}{\partial QoS_w^t} = \frac{1}{3} $); 2)
$Rep^t_w$ ($\frac{\partial G^t_w}{\partial Rep^t_w} = \frac{1}{3}
$); and 3) $B_w^t$ ($\frac{\partial G^t_w}{\partial B_w^t} =
\frac{1}{3 (n_t \mu_w - \epsilon)} $). Thus, the sign of the two
first partial derivatives is positive and the sign of
$\frac{\partial G^t_w}{\partial B_w^t}$ depends on the sign of the
maximum profit $n_t \mu_w - \epsilon$, so we are done. The mean
service fee depends on the strategy adopted by the service because
a competitive service receives higher fees $\mu_{w, CM}$ compared
to a cooperative one $\mu_{w, CO}$ ($\mu_{w, CM} > \mu_{w, CO}$).
The motivation behind this is that a competitive service for a
given task is the leader for that task while other cooperative
services are performing specific subtasks as asked by the leader.

\begin{proposition}\label{Complexity-Growth_Factor}
$G_w^t$ can be computed in time linear in the size of the window
$t$.
\end{proposition}

\begin{proof}
As shown in the second part of Equation \ref{eq:growthfactor}, the
function $QoS_w^t$ can be computed in time linear in the number of
tasks, which in turn is linear in the size of the window time.
Since $B_w^t$ is constant, the result follows from Proposition
\ref{Complexity-Rep}. $\Box$
\end{proof}

The above explained parameters and other additional parameters
which will be used in the rest of this chapter are listed and self
explained in Table \ref{t4:Preliminaries}.

\begin{table}
\centering
\caption{List of abbreviations.}
\begin{tabular}{|c|c||c|c|}
\hline
\textbf{Notation} & \textbf{Definition} & \textbf{Notation} & \textbf{Definition}\\
\hline\hline
$T_{QoS}^r$    & Required task QoS for the task $r$   & $T_{QoS}^t$      & Mean required task QoS during $t$\\
$QoS_w^r$      & Service $w$ QoS for the task $r$ & $QoS_w^t$        & $w$ QoS during the window time $t$\\
$Reward_w^r$   & Reward obtained by $w$ w.r.t. $r$    & $Reward_w^t$ & Reward to update the reputation\\
$Rep^t_w$      & Reputation assigned for $w$          & $B_w^t$          & Budget associated to service $w$ \\
$G^t_w$        & Growth factor of $w$ during $t$      & $\epsilon$       & Community membership fee\\
$\pi_{w,CM}^t$ & Competition payoff of $w$            & $\pi_{w,CO}^t$   & Cooperation payoff of $w$\\
$p_{w,CM}^t$   & Competition probability of $w$       & $p_{w,CO}^t$     & Cooperation probability of $w$\\
$COF_w^t$      & Cooperation fee of $w$               & $\beta_{w}$      & Profit of $w$\\
$\mu_{w, CM}$  & Mean service fee for competing $w$   & $\mu_{w, CO}$    & Mean service fee for cooperating $w$\\
$\tau_w^t$     & Coopetitive threshold of $w$         & $P_w^t$          & Probability of competing for $w$\\
$U_w^t$        & Utility of $w$                       & $E^t_w$          & Expected number of tasks\\

\hline
\end{tabular}
\label{t4:Preliminaries}
\end{table}

\section{Service Interactive Strategies}

The main goal of each individual service agent is to increase its
income (payoff). This income can be earned from tasks (or
requests) done by this service. In our model, services can decide
to compete to get a task from the master agent or to cooperate
with other services within a given collaborative network (the way
a collaborative network is set by a leader is based on the
cooperative services reputation and their QoS parameters that
should coincide with the required QoS). Therefore we define two
types of service strategies. First, when a service has higher
level of confidence based on its growth factor, it can compete to
get a task from the master and adopts the competitive strategy.
Second, when the service agent has a lower level of confidence
that it does not feel capable to compete, it waits for some other
services to cooperate with to perform some tasks \footnote{Through
the report, requests or tasks are supposed to be decomposable.},
and thus it adopts the cooperative strategy. Services estimate the
outcome of all the strategies and choose one of them accordingly.
This decision is not static but can change over time so service
agents can switch from one strategy to the other, and this dynamic
attitude is referred to as coopetition. The underlying decision
making process is presented in the next section.

\section{Theoretical Results}\label{Theoretical Results}

\subsection{Service Decision Making Procedure}
%In the previous section, we introduced web services' interaction
%strategies.
In this section, we explore in details the interaction strategies
and the outcome of each strategy in terms of services' utilities.
The main part of services' decision making procedure falls into
their growth factor analysis. In fact, the comparison of the
growth factor to a particular threshold is the main reason that
influences the service's decision to follow either competitive or
cooperative behavior. service agents initially compute this value
and compare it with their computed threshold. Generally the main
challenge is the threshold computation and we cope with this issue
in the rest of this section. We additionally use the obtained
results in the implemented environment and analyze their
effectiveness on services' strategic decision making procedures.

Figure \ref{decisionMakingProcedure} shows the decision making
process that is followed by a typical service. In case the service
agent is ready to compete, there is a chance that it bids for a
task if it has the required capabilities to accomplish that task,
or stays silent and returns to the cooperative status. But in case
the service agent is willing to cooperate, it has to wait for a
cooperation opportunity that could be triggered by another service
agent that competed and obtained the task, so both services will
be part of the same collaborative network. In the decision making
process presented in Figure \ref{decisionMakingProcedure}, we
assume that the competing service might get the task (denoted as
$Bid/obtainedTask$) or not in case of being rejected by the master
agent, or do not even bid for the task (denoted as
$Silent/rejectedTask$). For simplicity reasons and without loss of
generality, we group the two cases of $Bids$ and $obtainedTask$
together as well as $Silent$ and $rejectedTask$. The rational
behind this aggregation is the fact that our main concentration is
services' status (competitive or cooperative) over different
decision making rounds, which could be caused by internal factors
(the services) or by the external factor (the master agent).

\begin{figure}[h]
\centering
\includegraphics[scale=0.55]{Figures/decision1++.eps}
\caption{Decision making process over competitive and cooperative
strategies.} \label{decisionMakingProcedure}
\end{figure}

Consider a service $w$ that is willing to compete for the period
of time $t$ (that means the computed growth factor is more than
the analyzed threshold). This service can estimate the expected
payoff associated to this decision, called \emph{competition
payoff}. Equation \ref{CMPff} computes this expected payoff for
the competing service $w$ ($\pi_{w,CM}^t$) considering the
$Bid/obtainedTask$ probability of $p_{w,CM}^t$ and
$Silent/rejectedTask$ probability of $1-p_{w,CM}^t$.

\begin{equation}\label{CMPff}
\pi_{w,CM}^t=p_{w,CM}^t(\mu_{w, CM}E^t_w-COF_w^t
E^t_w-\epsilon)+(1-p_{w,CM}^t)(-\epsilon)
\end{equation}

In Equation \ref{CMPff}, $E^t_w$ is the number of tasks that $w$
expects during the window time $t$, and $\mu_{w, CM}$ is the mean
service fee that is assigned by the master agent to $w$. This
means that a competing service directly obtains this fee from the
master agent. Moreover, the competing service $w$ expects a
cooperation fee ($COF_w^t$) that it gives to its collaborators in
case $w$ needs to cooperate with other services (cooperative
service agents in its collaboration network). In any case, the
competing or cooperating service agent pays a fixed amount of
membership ($\epsilon$) to the master agent, which plays the role
of the community's coordinator. This fee would be taken into
account when a service decides to leave to a cheaper community or
act alone. But to concentrate on the main concerns of this paper,
we skip these small details.\\\\
%
From Equation \ref{CMPff}, the following proposition is
straightforward.
